---
title: "Problem Set Assignment No. 1"
author: "Aria Muchhal"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: default
---

```{r SetupChunk, echo = TRUE}

## Keep this line always
knitr::opts_chunk$set(echo = TRUE,
                      collapse = TRUE,
                      warning = FALSE, message = FALSE,
                      fig.align = 'center')


```

## 1. Setup your document

### A. Load Libraries

```{r q1A}
library(knitr)
library(wooldridge)
wage2=wooldridge::wage2
```

*In the R videos on MediaSpace, you are asked to install multiple packages. Name three of them here.*
rmarkdown, dplyr, and magrittr.

### B. Show installed packages

```{r q1B}
installed.packages()[,c(1,3)]
```

Write your answer to 1B here if requested.

## 2. Standard Error of the Mean

### A. Subset wage2

```{r q2A}
wage2 = wage2[wage2$educ==16,1]
NROW(wage2)
```
*How many rows does wage2 have once you have subset it? Use NROW to see.*
150

### B. Calculate the sample mean and sample variance for wage in our sample of wages
```{r q2B}
meanCollege=mean(wage2)
varCollege=var(wage2)
nCollege=NROW(wage2)
print(meanCollege)
print(varCollege)
print(nCollege)
```
*What are the units of the wage column? Hint: you can learn about wooldridge datasets by using the help for that data ?wage2*
monthly earnings

### C. Calculate the standard error of the mean
```{r q2C}
semCollege = sqrt(varCollege/nCollege)
print(semCollege)
```

## 3. Hypothesis testing: population mean

### A. Hypothesis testing H0 : µwage = 1, 200

```{r q3A}
dfn=nCollege-1
tcrit=qt(.975,dfn)
ci=(meanCollege-(tcrit*semCollege))
ci2=(meanCollege+(tcrit*semCollege))
print (ci)
print(ci2)
```

*What is your answer to this question, in your own words, and why?*
I would say no, because 1200 is not within the 95% confidence interval of (1045.71, 1171).
