---
title: "Problem Set Assignment No. 3"
author: "Aria Muchhal"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: default
editor_options: 
  markdown: 
    wrap: sentence
---

```{r SetupChunk, echo = TRUE}

## Keep this line always
knitr::opts_chunk$set(echo = TRUE,
                      collapse = TRUE,
                      warning = FALSE, message = FALSE,
                      fig.align = 'center')


```

## 0.

### Task A.

```{r q0A}
library(knitr)
library(tinytex)
library(wooldridge)
library(lmtest)
library(sandwich)
library(AER)
library(fixest)
library(modelsummary)

```

## 1.

### Task A.

```{r t1A}
data(CASchools)
table(CASchools$county)
```

### Question A.

*a. How many observations are in the data?*

420. 

*b. We are interested in test scores. Which variable(s) in CASchools would be our outcome of interest?*

read and math

*c. Using the output from the table(...) command, what county has the most observations in the data?*

Sonoma.

### Task B. Data Cleaning

(a) 

```{r t1b}
studentTeacherRatio=CASchools$students/CASchools$teachers
CASchools$studentTeacherRatio<-studentTeacherRatio
```

(b) 

```{r}
bigCounties=CASchools[CASchools$county=='Sonoma'|CASchools$county=='Los Angeles'|CASchools$county=='Kern', TRUE]
```

(c) 

```{r}
A=CASchools[CASchools$county=='Sonoma'|CASchools$county=='Los Angeles'|CASchools$county=='Kern', c('district', 'school', 'county', 'grades', 'students', 'teachers', 'calworks', 'lunch', 'computer', 'expenditure', 'income', 'english', 'read', 'math', 'studentTeacherRatio')]
```

(d) 

```{r}
plot(A$studentTeacherRatio, A$math, col="green")
plot(A$studentTeacherRatio, A$read)

```

(e) 

```{r}
plot(A$income, A$math, col="blue")
plot(A$income, A$read, col="blue")
```

(f) 

```{r}
plot(A$studentTeacherRatio, A$income, col=as.factor(A$county))
```

### Question B. Plot the relationship

*a. Using your first two plots, does there appear to be a relationship between higher student:teacher ratios and math scores? What about reading scores?*

One can see a slight negative correlation between math and reading scores and student:teacher ratios.
Essentially, higher ratios appear to be associated with lower scores.

*b. Using your last two plots, does there appear to be a relationship between higher income and math scores? What about reading scores?*

One can see a very strong positive correlation between math and reading scores and income.

*c. Using the final plot, does it appear that some counties have higher income or higher student:teacher ratios (or both)?*

While some counties are more dispersed than others, it does seem like 1 county has a higher income and lower student to teacher ratio, another has a higher income and higher student to teacher ratio, and the last has a lower income and dispersed student to teacher ratio.

### Task C

(a) 

```{r}
coeftest(lm(read ~ calworks, A), vcov = vcovHC, "HC1")
```

### Question C

*a. What is the coefficient on calworks and what does it mean? Note that calworks is in percentage points (you can see the range using range(CASchools\$calworks))*

-1.1971, so one additional point in a school's reading scores is associated with a 1.1971% decrease in calworks.

*b. What potential omitted variables might bias this coefficient? That is, is there something unobserved correlated with calworks that might also be correlated with read?*

Calworks would definitely be associated with income and lunch, and it is likely that both would be correlated with read.

### Task D

(a) 

```{r}
coeftest(lm(read ~ calworks+as.factor(county), A), vcov = vcovHC, "HC1")
```

### Question D

*a. What is the new coefficient on calworks? Is it larger or smaller?*

4.895e-06+2.2e-16=4.895e-6, which is smaller.

*b. What is the base county level?*

653.11817

*c. What is the expected reading score for an observation in Sonoma County, at a schools with a calworks value of 25%?*

653.11817+20.94302+(2.2e-16+1.180e-06)\*25=674.061219

### Task E

(a) 

```{r}
uhat1=residuals(lm(read ~ calworks+as.factor(county), A), na.rm=FALSE)
A$uhat1<-uhat1
uhat2=uhat1^2
A$uhat2<-uhat2
```

(b) Run the appropriate regression (see slide 108 of 02-Multivariate Regression) and show the results

```{r}
appropReg<-lm(uhat2 ~ calworks+as.factor(county), A)
summary(appropReg)
```

### Question E

*a. What is the relevant output of the regression summary for our Breusch-Pagan Test H0: "No Heteroskedasticity present"*

The p-value of the F statistic from this test.
If it's small, we reject the H0--homoskedasticity--which would indicate heteroskedasticity.

*b. What is your interpretation of the results? Should we be using heteroskedastic errors?*

The p-value is small: .008246, so we can reject the H0 and feel comfortable using heteroskedastic errors.

### Task F

```{r}
bptest(read ~ calworks+as.factor(county), data=A)
```

### Question F

*a. What is the interpretation of the result from this version of the Breusch-Pagan test?* The pvalue is quite low, 0.009679, so we should reject the H0 of "No Heteroskedasticity present" and use HC robust errors.

*b. Is it the same (or very close) to our results from Task 1.E?* It is not the same, but it is quite close.

### Task G

```{r}
coeftest(lm(read ~ calworks+english+calworks*english+as.factor(county), A), vcov = vcovHC, "HC1")

```

### Question G

*a. What is the effect of an increase in calworks by one unit for a school with 0% english learners (english=0)?*

     663.6423532-0.7711882=662.871165 increase in reading score, with all else held equal.

*b.What is the effect of an increase in calworks by one unit for a school with 40% english learners?*

    663.6423532-0.7711882-(0.8269718*.4)+(0.0158432*.4)=662.546714 increase in reading score, with all else held equal.

*c.What is the formula you used to determine dRead/dCalworks ? Hint: it includes the variable english. Write it using LaTeX.*

$\frac{dRead}{dCalworks}=\beta_0+\beta_{calworks}*calworks+\beta_{english}*english+\beta_{calworks-and-english}*(calworks-and-english)$


## 2. Last Question

I spent 9 continuous hours on this problem set, but it took around 2 days with breaks.
